%==============================================================================
\section{Introduction}
\label{sec:introduction}

Two-dimensional echocardiography is the most widely used cardiac imaging
modality worldwide: it is portable, inexpensive, non-ionising, and provides
real-time visualisation of cardiac structure and function. Its core
quantitative output --- left ventricular ejection fraction (LVEF), computed
from end-diastolic and end-systolic volumes via the biplane Simpson's method
\citep{lang2015} --- drives a large fraction of clinical cardiac decisions,
from diagnosis of heart failure to monitoring of cardiotoxic chemotherapy.
The clinical importance of LVEF is matched by the operator-dependent
nature of its measurement: published inter-observer LVEF variability is
approximately $7$--$10\%$ \citep{leclerc2019}, which directly limits the
sensitivity of LVEF as a clinical decision boundary.

Automated segmentation of the left ventricular endocardium (LV-endo),
epicardium (LV-epi), and left atrium (LA) from two-dimensional apical
2-chamber (2CH) and 4-chamber (4CH) views would, in principle, eliminate
this operator-dependent variability and provide reproducible volumetric
measurements. The technical challenge is substantial: ultrasound suffers
from low signal-to-noise ratio, speckle noise, low boundary contrast,
acoustic shadowing, and view-dependent geometric foreshortening.

\paragraph{Architectural pluralism.}
Three paradigms have emerged as competitive solutions in this space:
\begin{itemize}[leftmargin=*,nosep]
\item \textbf{Pure convolutional networks.} U-Net
      \citep{ronneberger2015} and its descendants (residual encoders, SE
      blocks, attention gates, deep supervision)
      \citep{milletari2016,oktay2018} remain the dominant architectural
      family, with nnU-Net \citep{isensee2021} serving as a strong
      auto-configured baseline.
\item \textbf{Hybrid CNN--Transformer networks.} TransUNet
      \citep{chen2021transunet} and similar hybrids combine a
      convolutional encoder with a Vision Transformer bottleneck,
      explicitly trading off local detail (CNN) against global context
      (self-attention).
\item \textbf{Pure Transformer networks.} Swin-UNet \citep{cao2022}
      replaces the convolutional backbone entirely with windowed
      self-attention, demonstrating that pure attention can match or
      exceed CNN performance on a range of medical tasks.
\end{itemize}

\paragraph{The benchmarking gap.}
Despite this architectural pluralism, the literature offers no
systematic, like-for-like comparison of these three paradigms on a single
shared evaluation pipeline. Each paper introducing a new architecture
typically compares against a small number of baselines under different
training protocols, augmentations, learning-rate schedules, and
evaluation metrics. As a result it is difficult for clinicians or
practitioners to answer a basic question: \emph{which architectural
family should I use for echocardiographic segmentation, and what are
the trade-offs?}

This paper provides that comparison on the CAMUS dataset
\citep{leclerc2019}, the standard public 2D echocardiographic
segmentation benchmark, with 500 patients across two views and two
cardiac phases. We train nine representative architectures spanning the
three paradigms above using one identical protocol --- same optimiser,
loss, schedule, augmentation, mixed-precision settings, and early-stopping
criterion --- and evaluate them on a single shared pipeline reporting
geometric, boundary, clinical, and efficiency metrics simultaneously.

\paragraph{Contributions.}
\begin{itemize}[leftmargin=*,nosep]
\item We present the largest unified benchmark of deep segmentation
      architectures on CAMUS to date: nine architectures across three
      paradigms, trained and evaluated under a single shared protocol.

\item We report \emph{multi-dimensional} evaluation, going beyond Dice
      to include HD95 in millimetres (with per-sample NIfTI pixel
      spacing), ASSD, biplane Simpson's EF MAE and Pearson $r$, ED/ES
      stratified metrics, robustness across CAMUS image-quality grades,
      and efficiency benchmarks (parameters, FLOPs, latency, peak
      memory). The full results matrix and per-patient outputs are
      released.

\item We apply pairwise Wilcoxon signed-rank tests with Bonferroni
      correction to identify which architectural differences are
      statistically significant, and produce a critical-difference
      diagram showing the top-tier band of equivalent models.

\item Our best architecture, TransUNet, reaches Dice $= 0.9122$,
      HD95 $= 1.82$~mm; nnU-Net reaches Dice $= 0.9099$, HD95 $= 1.93$~mm
      with $6\times$ fewer parameters and \PLACE{$\sim 5\times$} faster
      inference. The HD95 of approximately $1.9$~mm represents a
      substantial improvement over the originally published nnU-Net
      baseline ($\approx 4.3$~mm) \citep{leclerc2019} and demonstrates
      the importance of training-protocol modernisation for fair
      cross-architecture comparison.

\item We release a fully reproducible benchmark framework: training,
      evaluation, statistical analysis, and figure generation code.
\end{itemize}

\paragraph{Outline.}
Section~\ref{sec:related} reviews segmentation architectures and the
CAMUS benchmark. Section~\ref{sec:methods} describes the nine
architectures, the unified training protocol, and the evaluation
pipeline. Section~\ref{sec:results} reports the experimental results.
Section~\ref{sec:discussion} discusses architectural design choices,
trade-offs, and limitations. Section~\ref{sec:conclusion} concludes.
