%==============================================================================
\section{Related Work}
\label{sec:related}

\subsection{Convolutional segmentation networks}
\label{sec:related-cnn}

U-Net \citep{ronneberger2015} established the encoder--decoder
architecture with skip connections that remains the dominant template for
medical image segmentation. Subsequent work has refined the encoder
(residual blocks \citep{he2016}, ResNet
\citep{he2016}, EfficientNet), the decoder
(transposed convolution vs.\ bilinear upsampling, dense connections
\citep{huang2017}), and the skip mechanism
(attention gates \citep{oktay2018}, dense skips). Squeeze-and-excitation
blocks \citep{hu2018} provide channel-wise recalibration. Atrous spatial
pyramid pooling (ASPP) in DeepLabV3+ \citep{chen2018encoder} replaces the
deep encoder with dilated convolutions to enlarge the receptive field
without down-sampling. Feature Pyramid Networks
\citep{lin2017} aggregate features across multiple scales for dense
prediction.

The nnU-Net framework \citep{isensee2021} integrates these architectural
choices with automatic data-dependent hyperparameter selection (target
spacing, patch size, batch size, augmentation pipeline) and has set
strong baselines on numerous medical benchmarks including CAMUS.

\subsection{Transformer-based segmentation networks}
\label{sec:related-trans}

Self-attention's quadratic complexity made global attention impractical
for high-resolution images until Vision Transformers
\citep{dosovitskiy2021} demonstrated patch-based tokenisation. TransUNet
\citep{chen2021transunet} was the first work to bring this design to
medical segmentation, combining a CNN encoder (typically ResNetV2-BiT
\citep{kolesnikov2020}) with a ViT-B/16 bottleneck and a cascaded
upsampler decoder. The hybrid design retains the inductive bias of
convolutions at high resolution while granting global context at the
bottleneck.

Swin Transformer \citep{liu2021swin} introduced windowed self-attention
with shifted windows, restoring linear complexity in image size. Swin-UNet
\citep{cao2022} adapts this to segmentation with a fully Transformer-based
U-Net. SwinV2 \citep{liu2022swinv2} and Swin-MAE
\citep{baochen2023swinmae} further refine pretraining strategies.

\subsection{Echocardiographic segmentation and CAMUS}
\label{sec:related-camus}

The CAMUS dataset \citep{leclerc2019} (\emph{Cardiac Acquisitions for
Multi-structure Ultrasound Segmentation}) provides 2D apical-2-chamber and
apical-4-chamber acquisitions from 500 patients, each annotated by an
expert echocardiographer at end-diastole and end-systole with four
classes (background, LV-endo, LV-epi, LA). The official split assigns
$400$ patients to training, $50$ to validation, and $50$ to test, with
test annotations held out by the challenge organisers. Image quality is
graded into \emph{Good}, \emph{Medium}, and \emph{Poor} categories by
the same experts.

The original CAMUS paper reported nnU-Net achieving
HD $\approx 4.3$~mm on LV-endo and $\sim$$9\%$ EF MAE.
Subsequent work \citep{painchaud2022, smistad2021}
has reported incremental improvements using deeper architectures, custom
losses, and view-aware training. Several papers have additionally
explored sequence-aware models on CAMUS half-cycle annotations
\citep{painchaud2022}, but these are out of scope for the present
benchmark.

EchoNet-Dynamic \citep{ouyang2020} provides a complementary large-scale
video dataset for EF regression; we focus on CAMUS as the standard
benchmark for \emph{segmentation-derived} EF, but our findings about
architectural trade-offs are likely to transfer.

\subsection{What is missing}
\label{sec:related-gap}

A small number of recent works compare a handful of architectures on
CAMUS in their experimental sections, but none provide:
\begin{itemize}[leftmargin=*,nosep]
\item A unified training protocol across CNN, hybrid, and Transformer
      paradigms (papers typically inherit each architecture's own
      published hyperparameters, confounding architecture with
      protocol).
\item Per-class HD95 in millimetres with per-sample NIfTI spacing
      (most papers report Dice only, or use pixel-space HD95).
\item ED- versus ES-stratified reporting (essential for biplane
      Simpson's EF analysis).
\item Robustness analysis by CAMUS image-quality grade.
\item Pairwise Wilcoxon significance with Bonferroni correction and a
      critical-difference diagram.
\item Released model checkpoints and per-patient predictions for
      independent reanalysis.
\end{itemize}
This paper aims to fill each of these gaps.
