%==============================================================================
\section{Discussion}
\label{sec:discussion}

\subsection{Which architectural family wins on CAMUS?}
\label{sec:disc-families}

The top tier of our benchmark is mixed across paradigms: a hybrid
CNN--Transformer (TransUNet) leads on bulk Dice and HD95, but a pure CNN
(nnU-Net) leads on EF MAE \emph{and} on parameter efficiency. The
difference between them on mean Dice is $0.0023$ --- below the
Bonferroni-corrected statistical-significance threshold. A practitioner
choosing between these two architectures should be guided not by Dice
but by deployment constraints (nnU-Net needs $6.5\times$ fewer parameters
and is faster) or by the clinical objective (nnU-Net is more accurate at
EF and tighter at Bland--Altman bias).

The pure-Transformer family (Swin-UNet) is competitive on bulk Dice
($0.8878$) but visibly worse on EF (MAE $10.46\%$, $r = 0.574$). This
finding is consistent with reports that pure-attention models, despite
strong overall overlap, can produce boundary contours that are subtly
biased \citep{chen2021transunet} --- an effect that compounds when
contours feed Simpson's biplane volumetric integration.

The bottom of the table is occupied by FPN-UNet, DeepLabV3+, and
DenseContextU-Net. All three were designed for natural-image semantic
segmentation (FPN/ASPP) or for image classification (dense connectivity)
respectively. Their underperformance on CAMUS is consistent with the
hypothesis that medical-segmentation-specific designs (U-Net's skip
hierarchy, nnU-Net's training recipe) matter more on this dataset than
generic semantic-segmentation prowess.

\subsection{Comparison to the published CAMUS baselines}
\label{sec:disc-leaderboard}

The original CAMUS paper \citep{leclerc2019} reported nnU-Net HD95
$\approx 4.3$~mm and EF MAE $\sim$$9\%$. Our reproduced nnU-Net under the
unified protocol reaches HD95 $1.93$~mm and EF MAE $5.72\%$ --- roughly
twice as accurate on the boundary axis and substantially better on EF.

This gap is \emph{not} an architectural improvement: nnU-Net's
architecture is unchanged. The improvement is attributable to three
training-protocol upgrades made standard since 2019:
\begin{enumerate}[leftmargin=*,nosep]
\item Cosine-with-warm-restarts learning-rate schedule with
      $10$-epoch warmup, in place of the original ReduceLROnPlateau.
\item Mixed-precision training (FP16 AMP), which allows larger effective
      batch sizes and smoother loss landscapes.
\item Composite loss with an explicit boundary term
      \citep{kervadec2019}, which directly penalises the
      Hausdorff-relevant boundary error.
\end{enumerate}
This is a useful baseline calibration point for the field: cross-paper
comparisons of CAMUS results from before $\sim 2021$ to after should be
read with the understanding that protocol changes alone produce
sub-millimetre HD95 improvements at constant architecture.

\subsection{Design choices that matter}
\label{sec:disc-design}

Aggregating across our nine architectures, several design choices
correlate with strong performance:

\begin{itemize}[leftmargin=*,nosep]
\item \textbf{Pretrained encoders} (TransUNet, UNet-ResNet, Swin-UNet,
      FPN-UNet, DeepLabV3+). The two highest-Dice non-nnU-Net models use
      pretrained encoders. However, in a parameter-matched
      analysis (reported in the companion SSM study), this advantage
      shrinks substantially.
\item \textbf{Deep skip connections} (every U-Net-shaped architecture
      except DeepLabV3+). DeepLabV3+'s lightweight decoder, while
      efficient, appears to limit high-resolution boundary precision on
      CAMUS.
\item \textbf{Boundary-aware loss} (used uniformly in our protocol).
      This is a protocol choice rather than an architecture choice, but
      it appears to be among the largest single contributors to the
      HD95 improvement over published CAMUS baselines.
\item \textbf{Auto-configuration} (nnU-Net's distinguishing feature).
      The nnU-Net framework produces a competitive model on this dataset
      with $\sim$$15$~M parameters --- substantially smaller than any
      pretrained-encoder architecture --- by selecting patch size, batch
      size, and augmentation strategy from data statistics.
\end{itemize}

\subsection{Clinical implications}
\label{sec:disc-clinical}

For deployment in echocardiographic LVEF assessment, the choice between
the four top-tier architectures should be driven by three criteria:
(i) clinical accuracy target, (ii) hardware constraints, and (iii)
inference-time latency budget.
\begin{itemize}[leftmargin=*,nosep]
\item For \emph{maximum reproducibility and tightest EF agreement},
      nnU-Net is the recommended choice (lowest EF MAE, smallest
      Bland--Altman bias).
\item For \emph{maximum boundary accuracy} at deployment-irrelevant cost,
      TransUNet is the highest-HD95 model.
\item For \emph{edge / portable ultrasound deployment} where
      parameters and latency dominate, nnU-Net at $15.6$~M parameters
      remains the best trade-off; no smaller architecture in our
      benchmark reaches comparable Dice.
\end{itemize}
None of the four top-tier models reach below-$5\%$ EF MAE in this study,
indicating that human-level cardiologist agreement is still a target
boundary for the field rather than a solved problem.

\subsection{Limitations and future work}
\label{sec:disc-limits}

We acknowledge five limitations.

\begin{enumerate}[leftmargin=*,nosep]
\item \textbf{Single-dataset.} All evaluation is on CAMUS. A cross-dataset
      transfer evaluation (\emph{e.g.,} EchoNet-Dynamic
      \citep{ouyang2020}) would strengthen the generality claims and is
      planned for future work.
\item \textbf{2D only.} CAMUS provides 2D apical views; 3D
      echocardiography is increasingly common in modern practice and
      requires architectures and pretraining recipes outside the scope
      of this benchmark.
\item \textbf{Test set size.} The CAMUS official test set is 50 patients
      ($200$ ED+ES frames). All point estimates are reported with
      bootstrap $95\%$ confidence intervals to reflect this.
\item \textbf{No SSM models.} The recent wave of selective state space
      models (Mamba family) is the subject of a separate, dedicated study
      (a companion study), and is deliberately excluded
      from the present benchmark to maintain clean three-paradigm
      framing.
\item \textbf{Single training protocol.} A single shared protocol is the
      goal of this benchmark, but some architectures may benefit from
      protocol-specific tuning (\emph{e.g.,} larger batch sizes for
      pretrained Transformers). Per-architecture hyperparameter searches
      could shift rankings within the top-tier band.
\end{enumerate}

\subsection{Take-home messages}
\label{sec:disc-takehome}

\begin{enumerate}[leftmargin=*,nosep]
\item No single architectural family dominates CAMUS. Hybrid
      (TransUNet) leads on bulk Dice; pure CNN (nnU-Net) leads on EF
      accuracy and parameter efficiency. Pure Transformer (Swin-UNet) is
      competitive on Dice but degraded on EF.
\item Training-protocol modernisation (cosine schedule, AMP, boundary
      loss) accounts for most of the $\sim$$2\times$ improvement in
      HD95 over the original CAMUS baselines, independent of
      architecture.
\item LV-epicardium remains the hardest of the three foreground classes
      across every architectural family.
\item The accuracy--parameter Pareto frontier is dominated by nnU-Net at
      the efficient end and TransUNet at the high-accuracy end.
\end{enumerate}
