%==============================================================================
\section{Conclusion}
\label{sec:conclusion}

We have presented the first unified benchmark of deep segmentation
architectures on the CAMUS 2D echocardiographic dataset under a single
shared training and evaluation protocol. Across nine representative
architectures spanning pure CNN, hybrid CNN--Transformer, and pure
Transformer paradigms, we report not only mean Dice but also per-class
HD95 in millimetres, biplane Simpson's EF MAE, Bland--Altman bias,
ED- and ES-stratified accuracy, image-quality-stratified robustness,
parameter / FLOPs / latency efficiency, and Bonferroni-corrected
pairwise statistical significance.

The hybrid TransUNet leads on bulk Dice ($0.9122$) and HD95
($1.82$~mm). The pure-CNN nnU-Net leads on EF accuracy
(MAE $5.72\%$, $r = 0.823$) and is $6.5\times$ more parameter-efficient.
The top four architectures (TransUNet, nnU-Net, UNet-ResNet, UNet-V1)
form a statistically equivalent top-tier band. The pure-Transformer
Swin-UNet is mid-tier on bulk Dice but degraded on EF, suggesting that
its contours are subtly biased in a manner that compounds through
Simpson's biplane integration.

The full-network HD95 of approximately $1.9$~mm represents a substantial
improvement over the originally reported CAMUS nnU-Net baseline of
$4.3$~mm. We attribute most of this improvement to training-protocol
modernisation (cosine annealing, mixed precision, boundary loss) rather
than to architectural changes. This observation is useful as a baseline
calibration point for cross-paper comparison of CAMUS results.

Three directions remain open. (i) Cross-dataset transfer to
EchoNet-Dynamic and ACDC would test whether the architectural ranking
holds beyond CAMUS. (ii) 3D echocardiography requires architectures and
pretraining recipes outside the scope of this benchmark. (iii) The
recent wave of selective state space models (Mamba, Mamba-2, VMamba) is
the subject of a separate, dedicated companion study
(a companion study).

We release the complete training, evaluation, statistical-analysis, and
figure-generation pipeline, together with the nine model checkpoints
and per-patient JSON / CSV outputs that produced this paper, to lower
the barrier to reproducing and extending these results.
